### 2022-06-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@990f852bfb5132c073bedfca4370007dad21b749 🚀

\vfill
Pathogen evolution is one of the least predictable components of disease
emergence, particularly in zoonotic reservoirs. Here, building on
principles established by the geographic mosaic theory of coevolution,
we develop a quantitative, spatially-explicit framework for mapping the
evolutionary risk of viral emergence. Driven by interest in diseases
like SARS, MERS, and COVID-19, we examine the global biogeography of
bat-origin betacoronaviruses, and apply our coevolutionary framework to
develop a map of emergence risk. We find that our framework helps
explain patterns like a unique pool of merbecoviruses in the Neotropics,
a recently-discovered lineage of divergent nobecoviruses in Madagascar,
and---most importantly---hotspots of diversification in southeast Asia,
sub-Saharan Africa, and the Middle East that correspond to the site of
previous zoonotic emergence events. Our framework may help identify
hotspots of future risk that have also been previously overlooked, like
west Africa and the Indian subcontinent, and may more broadly help
researchers understand how reservoir ecology shapes the evolution of
pandemic threats.

>>>>

\vfill
Pathogen evolution is one of the least predictable components of disease
emergence, particularly in nature. Here, building on principles
established by the geographic mosaic theory of coevolution, we develop a
quantitative, spatially-explicit framework for mapping the evolutionary
risk of viral emergence. Driven by interest in diseases like SARS, MERS,
and COVID-19, we examine the global biogeography of bat-origin
betacoronaviruses, and find that our coevolutionary framework generates
different inferences about risk than simply mapping hotspots of host
diversity. Further, our framework helps explain patterns like a unique
pool of merbecoviruses in the Neotropics, a recently-discovered lineage
of divergent nobecoviruses in Madagascar, and--most
importantly--hotspots of diversification in southeast Asia, sub-Saharan
Africa, and the Middle East that correspond to the site of previous
zoonotic emergence events. Our framework may help identify hotspots of
future risk that have also been previously overlooked, like west Africa
and the Indian subcontinent, and may more broadly help researchers
understand how host ecology shapes the evolution and diversity of
pandemic threats.

### 2022-06-29 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@990f852bfb5132c073bedfca4370007dad21b749 🚀

Pathogen evolution is one of the least predictable components of disease
emergence, particularly in zoonotic reservoirs. Here, building on
principles established by the geographic mosaic theory of coevolution,
we develop a quantitative, spatially-explicit framework for mapping the
evolutionary risk of viral emergence. Driven by interest in diseases
like SARS, MERS, and COVID-19, we examine the global biogeography of
bat-origin betacoronaviruses, and apply our coevolutionary framework to
develop a map of emergence risk. We find that our framework helps
explain patterns like a unique pool of merbecoviruses in the Neotropics,
a recently-discovered lineage of divergent nobecoviruses in Madagascar,
and---most importantly---hotspots of diversification in southeast Asia,
sub-Saharan Africa, and the Middle East that correspond to the site of
previous zoonotic emergence events. Our framework may help identify
hotspots of future risk that have also been previously overlooked, like
west Africa and the Indian subcontinent, and may more broadly help
researchers understand how reservoir ecology shapes the evolution of
pandemic threats.

>>>>

Pathogen evolution is one of the least predictable components of disease
emergence, particularly in nature. Here, building on principles
established by the geographic mosaic theory of coevolution, we develop a
quantitative, spatially-explicit framework for mapping the evolutionary
risk of viral emergence. Driven by interest in diseases like SARS, MERS,
and COVID-19, we examine the global biogeography of bat-origin
betacoronaviruses, and find that our coevolutionary framework generates
different inferences about risk than simply mapping hotspots of host
diversity. Further, our framework helps explain patterns like a unique
pool of merbecoviruses in the Neotropics, a recently-discovered lineage
of divergent nobecoviruses in Madagascar, and--most
importantly--hotspots of diversification in southeast Asia, sub-Saharan
Africa, and the Middle East that correspond to the site of previous
zoonotic emergence events. Our framework may help identify hotspots of
future risk that have also been previously overlooked, like west Africa
and the Indian subcontinent, and may more broadly help researchers
understand how host ecology shapes the evolution and diversity of
pandemic threats.
